---
title: "Hands_on_sessions_1_2_3"
author: "Janet Piñero (janet.pinero@upf.edu)"
date: "Last update: `r format(Sys.time(), '%d %B, %Y')`"      
output:
  html_document:
    toc: true
    fig_caption: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Pima indians diabetes dataset 

Load the Pima indians diabetes dataset (Smith et al, 1988) from the mlbench package. 

Find the description of the dataset and classify the attributes according to their type

```{r}
# install.packages("mlbench")
# install.packages('curl')
# install.packages("data.table")
# install.packages("dplyr")

library(mlbench)
library(data.table)
library(dplyr)
data("PimaIndiansDiabetes")

```

What is the average age of the people in the study?

```{r}
 

```

How many cases and controls are there in the study? 
```{r}
 

```

Are there missing values in the dataset? 

```{r}



```

Which variables/features include outliers? How did you detected them?

```{r}

```

Perform imputation on missing data. Justify your choices

```{r}

```


From the Pima indians dataset, and visualization hands on,  plot using ggplot2:
Do not use the colors by default from ggplot2!! Add the plot titles. 

The density plot of Pregnancies Grouped by outcome (in this case, outcome refers to presence of diabetes)
```{r}

```

The density plot of glucose by presence outcome, and next to it, the boxplot of glucose by outcome 
```{r}

```

The relationship between glucose and insulin levels colored by outcome
```{r}

```

The relationship between the  Skinfold thickness of the triceps and the BMI, using one panel for each outcome
```{r}

```


# 2. Worcester Heart Attack Study 

Load the whas.csv dataset in the data folder, from the Worcester Heart Attack Study (Saczynski et al, 2008).

```{r}


```


What is the average length of stay in the hospital	
```{r}

```

What is the average length of the follow up?

```{r}


```


Plot the distribution of the age by gender. Is there any trend? What is the best way to visualize this?** 
```{r}

```


# 3 the Heart Disease Dataset 
Load the heart_disease_dataset.csv dataset in the data folder, and examine the description of the variables in the file heart_disease_description.txt. 
Then, perform an EDA on the data
```{r}

```

Is this a balanced dataset? Can you do something to solve this?*
```{r}

```


Analyze the dataset and answer which variables seem more predictive of heart disease*
```{r}

```



# 4  


# session info {.unnumbered}

```{r, results='asis',  echo=FALSE, message=FALSE }
sessionInfo()
```
